\documentclass[12pt]{article}

%\pagestyle{empty}
\textwidth 6.5in
\textheight 9in
\oddsidemargin 0in
\topmargin 0 in
\headheight 0 in
\headsep 0in
%\footheight 0in
\footskip 0.5in
\parindent 0 in

\usepackage{calc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{etaremune}
\usepackage{multicol}
\usepackage{titlesec}
\hypersetup{colorlinks=true,linkcolor=blue,anchorcolor=blue,citecolor=blue,filecolor=blue,urlcolor=blue}

%\usepackage[letter, total={6.5in, 9in}]{geometry}
%\usepackage[letterpaper, margin=1in]{geometry}

\titleformat{\section}{\normalsize \bfseries}{\thesection}{0.5 em}{}
\setlength{\topmargin}{0 cm}
\titlespacing{\section}{0pt}{1 em}{0.5 em}

\titleformat{\subsection}{\normalsize \bfseries}{\thesection}{0.5 em}{}
\titlespacing{\subsection}{0pt}{0.5 em}{0.5 em}
\setlength{\topmargin}{0 cm}

\providecommand*\email[1]{\href{mailto:#1}{#1}}

\begin{document}
\begin{center}
{\bf \Large Curriculum Vitae}\\
{\bf Avinash Rustagi}
\end{center}

\newlength{\rcollength}\setlength{\rcollength}{2.93in}
\begin{tabular}[t]{@{}p{\textwidth-\rcollength}p{\rcollength}}
Postdoctoral Research Associate & Email: \email{rustagi.avinash@gmail.com} \\ 
School of Electrical and Computer Engineering & \url{http://avinashrustagi.github.io/} \\ 
Purdue University, West Lafayette, IN 47906 & \\
\end{tabular} 

\renewcommand{\labelitemii}{$\cdot$}
\section*{\underline{Professional Employment}}
\begin{itemize}
\itemsep-0.05em
\item July 2018 - Present: Postdoctoral Research Associate, School of Electrical and Computer Engineering, Purdue University.

\item Sep 2016 - July 2018: Postdoctoral Research Scholar, Department of Physics, North Carolina State University. 

\item May 2016 - Aug 2016: OPS Research Assistant, Department of Physics, University of Florida.
\end{itemize}

\section*{\underline{Education}}
\begin{itemize}
\itemsep-0.05em
\item{Ph.D. in Physics, University of Florida (2016).}
\item{M.Sc. in Physics, Indian Institute of Technology, Kanpur, India (2010).}
\item{B.Sc. in Physics, St. Stephens College, Delhi, India (2008).}
\end{itemize}

\section*{\underline{Research Interests}}
\begin{itemize}
\itemsep-0.05em
\item Quantum information processing with magnons, Quantum sensing of novel phases.
\item Spintronics, van der Waal Magnonics.
\item Many body theory: Correlations in electron-hole systems, Ultrafast carrier and lattice dynamics, Non-equilibrium Bethe-Salpeter equation.
\item Electronic structure calculations: Density Functional Theory.
\item Optical and Transport Properties: Magneto-optics, Semiclassical Transport, Terahertz generation, Coherent Phonons.   
\end{itemize}

\section*{\underline{Honors and Awards}}
%\subsection*{ North Carolina State University, Graduate School}
\begin{itemize}
\itemsep-0.1em
\item Notable poster-College of Science, NC State Postdoctoral Research Symposium (2018).
\item College of Liberal Arts and Sciences (CLAS) Dissertation Fellowship funded by Threadgill Scholarship Program, Univ. of Florida (2016).
\item $5^{th}$ International Symposium on Terahertz Nanoscience Student Award, Martinique (2014).
\item Certificate of Outstanding Achievement for Academic Excellence, Univ. of Florida (2010-2014).
\item Center for Condensed Matter Sciences Summer Fellowship, Univ. of Florida (2011).
\item General Proficiency Medal for Academic Excellence, Indian Institute of Technology, Kanpur (2008-2010).
\item Academic Excellence Award, Indian Institute of Technology, Kanpur (2009-2010).
\item University Gold Medal for First Rank in B.Sc. Physics, St. Stephens College, Delhi (2008).
\item The Sumitomo Corporation-St. Stephens College Scholarship (2005-2008).
\item Ramesh Goel Memorial and Tushar Nagia prizes for best student,  St. Stephens College, Delhi (2006-2007). 
\end{itemize}

\section*{\underline{Computational Skills}} 
\begin{itemize}
\itemsep-0.1em
\item Quantum computing programming - Qiskit.
\item Micromagnetic simulations - mumax and OOMMF.
\item Programming in C++ (Parallel programming using MPI and OpenMP), Python, MATLAB, FORTRAN, Mathematica, and Shell Script. 
\item Density Functional Theory - Quantum Espresso for electronic structure calculations.
\end{itemize}

\section*{\underline{Publications}}
\subsection*{Published}
\begin{etaremune}
\itemsep-0.05em
\item A. B. Solanki, S. Bogdanov, {\bf A. Rustagi}, N. R. Dilley, T. Shen, M. M. Rahman, W. Tong, P. Debashish, Z.Chen, J. Appenzellar, Y. P. Chen, V. M. Shalaev, and P. Upadhyaya - \textit{Electric field control of interaction between magnons and quantum defect spins} \href{https://arxiv.org/abs/2012.01497}{\bf arXiv 2012.01497 (2020)}.

\item {\bf Avinash Rustagi}, Iacopo Bertelli, Toeno van der Sar, and Pramey Upadhayaya - \textit{Sensing chiral magnetic noise via quantum impurity relaxometry} \href{https://link.aps.org/doi/10.1103/PhysRevB.102.220403}{\bf [Phys. Rev. B 102, 220403 (Rapid Communications) (2020)}.

\item R. L. Wilmington, H. Ardekani, {\bf A. Rustagi}, A. Bataller, A. F. Kemper, R. A. Younts, and K. Gundogdu - \textit{Fermi liquid theory sheds light on ``Hot" EHL in 1L-MoS$_2$} \href{https://arxiv.org/abs/2010.09609}{\bf [arXiv 2010.09609 (2020)]}.

\item J. Huang, Z. Wang, H. Pang, H. Wu, H. Cao, S. Mo, {\bf A. Rustagi}, A. F. Kemper, R. J. Birgeneau, M. Wang, and M. Yi - \textit{Flatband-Induced Itinerant Ferromagnetism in RbCo$_2$Se$_2$} {\bf [Under review (2020)]}

\item Alexander F. Kemper and {\bf Avinash Rustagi} - \textit{Observing coherences with time-resolved photoemission} \href{https://arxiv.org/abs/2005.08978}{\bf [arXiv 2005.08978 (2020)]}.

\item {\bf Avinash Rustagi}, Abhishek Solanki, Yaroslav Tserkovnyak, and Pramey Upadhyaya - \textit{Coupled spin-charge dynamics in magnetic van der Waals heterostructures} \href{https://link.aps.org/doi/10.1103/PhysRevB.102.094421}{\bf [Phys. Rev. B 102, 094421 (2020)]}.

\item Terry Y.T. Hung, {\bf A. Rustagi}, S. Zhang, P. Upadhyaya, and Z. Chen - \textit{Experimental observation of coupled valley and spin Hall effect in p-doped WSe$_2$ devices} \href{http://dx.doi.org/10.1002/inf2.12095} {\bf[InfoMat. 2020; 1-7 (2020)]}.

\item Y. Jiang, Z. Lu, J. Gigliotti, {\bf A. Rustagi}, L. Chen, C. Berger, W. A. de Heer, C. J. Stanton, D. Smirnov, and Z. Jiang - \textit{Valley and Zeeman Splittings in Multilayer Epitaxial Graphene Revealed by Circular Polarization Resolved Magneto-infrared Spectroscopy} \href{https://doi.org/10.1021/acs.nanolett.9b02505}{\bf [Nano Letters 2019, 19, 10, 7043-7049]}.

\item J. Kaiser, {\bf A. Rustagi}, K. Y. Camsari, J. Z. Sun, S. Datta, and P. Upadhyaya - \textit{Subnanosecond Fluctuations in Low-Barrier Magnets} \href{https://doi.org/10.1103/PhysRevApplied.12.054056}{\bf [Phys. Rev. Applied 12, 054056 (2019)]}.

    \item {\bf Avinash Rustagi}, and Alexander F. Kemper - \textit{Coherent Excitonic Quantum Beats in Time-Resolved Photoemission Measurements} \href{http://doi.org/10.1103/PhysRevB.99.125303}{\bf [Phys. Rev. B 99, 125303 (2019)]}.

    \item A. W. Bataller, R. Younts, {\bf A. Rustagi}, Y. Yu,  H. Ardekani, A. F. Kemper, L. Cao, and K. Gundogdu - \textit{Dense Electron-Hole Plasma Formation and Ultra-Long Charge Lifetime in Monolayer MoS$_2$ via Material Tuning }\href{http://doi.org/10.1021/acs.nanolett.8b04408}{\bf  [Nano Letters 2019, 19, 1104-1111]}.

    \item O. Abdurazakov, D. Nevola, {\bf A. Rustagi}, J. K. Freericks, D. B. Dougherty, and A. F. Kemper - \textit{Non-equilibrium Electron Dynamics in Pump-Probe Spectroscopy: Role of excited phonon populations} \href{http://doi.org/10.1103/PhysRevB.98.245110}{\bf  [Phys. Rev. B 98, 245110 (2018)]}.

    \item {\bf Avinash Rustagi}, and Alexander F. Kemper - \textit{Photoemission signature of excitons} \href{https://link.aps.org/doi/10.1103/PhysRevB.97.235310}{\bf [Phys. Rev. B 97, 235310 (2018)]}.

	\item {\bf Avinash Rustagi}, and Alexander F. Kemper - \textit{Theoretical phase diagram for the room temperature Electron-Hole Liquid in photo-excited quasi-2D monolayer MoS$_2$} \href{http://pubs.acs.org/doi/abs/10.1021/acs.nanolett.7b04377} {\bf [Nano Letters 2018 18 (1), 455-459]}.

	\item Kunie Ishioka, {\bf Avinash Rustagi}, Andreas Beyer, Wolfgang Stolz, Kerstin Volz, Ulrich Hoefer, Hrvoje Petek, and Christopher J. Stanton - \textit{Sub-picosecond acoustic pulses generated at buried GaP/Si interfaces} \href{http://aip.scitation.org/doi/full/10.1063/1.4997913}{\bf  [Appl. Phys. Lett. 111, 062105(2017)]}.
 
	\item Kevin L. Pollock, Hoang Q. Doan, {\bf Avinash Rustagi}, Christopher J. Stanton, and Tanja Cuk - \textit{Detecting the Photoexcited Carrier Distribution Across GaAs/Transition Metal Oxide Interfaces by Coherent Longitudinal Acoustic Phonons} \href{http://pubs.acs.org/doi/abs/10.1021/acs.jpclett.6b02835} {\bf [J. Phys. Chem. Lett., 2017, 8, pp 922–928]}.

       \item Kunie Ishioka, {\bf Avinash Rustagi}, Ulrich Hofer, Hrvoje Petek, Christopher J. Stanton - \textit{Intrinsic coherent acoustic phonons in the indirect band gap semiconductors Si and GaP} \href{http://journals.aps.org/prb/abstract/10.1103/PhysRevB.95.035205} {\bf [Phys. Rev. B 95, 035205 (2017)]}.

			\item {\bf A. Rustagi} and C. J. Stanton - \textit{Terahertz radiation from accelerating charge carriers in graphene under ultrafast photoexcitation} \href{https://journals.aps.org/prb/abstract/10.1103/PhysRevB.94.195207} {\bf [Phys. Rev. B 94, 195207 (2016)]}.

      \item K. Ishioka, K. Brixius, A. Beyer, {\bf A. Rustagi}, C. J. Stanton, W. Stolz, K. Volz, U. Hofer and H. Petek - \textit{Coherent phonon spectroscopy characterization of electronic bands at buried semiconductor heterointerfaces} \href{http://scitation.aip.org/content/aip/journal/apl/108/5/10.1063/1.4941397} {\bf [Appl. Phys. Lett. 108, 051607 (2016)]}.

     \item K. Ishioka, K. Brixius, U. H{\"o}fer, {\bf A. Rustagi} , E. Thatcher, C. J. Stanton and H. Petek - \textit{Dynamically Coupled Plasmon-Phonon Modes in GaP; an Indirect-Gap, Polar Semiconductor} \href{http://journals.aps.org/prb/abstract/10.1103/PhysRevB.92.205203}{\bf [Phys. Rev. B 92, 205203 (2015)]}.

     \item {\bf A. Rustagi} and C. J. Stanton - \textit{Hot-electron noise properties of graphene-like systems} \href{http://journals.aps.org/prb/abstract/10.1103/PhysRevB.90.245424}{\bf [Phys. Rev. B 90, 245424 (2014)]}.

    \item L. G. Booshehri, C. H. Mielke, D. G. Rickel, S. A. Crooker, Q. Zhang, L. Ren, E. H. Hároz, {\bf A. Rustagi}, C. J. Stanton, Z. Jin, Z. Sun, Z. Yan, J. M. Tour, and J. Kono - \textit{Circular polarization dependent cyclotron resonance in large-area graphene in ultrahigh magnetic fields} \href{http://journals.aps.org/prb/abstract/10.1103/PhysRevB.85.205407} {\bf [Phys. Rev. B 85, 205407 (2012)]}.

\end{etaremune}

\vspace{-0.1in}
\subsection*{In preparation}
\begin{etaremune}
\itemsep-0.05em

\item Mohammad Mushfiqur Rahman, {\bf Avinash Rustagi}, Yaroslav Tserkovnyak, and Pramey Upadhyaya - \textit{Electrical generation and tunable routing of domain wall spin waves in van der Waals antiferromagnets} \href{}{\bf (2020)}.

 \item {\bf Avinash Rustagi}, Shivam Kajale, Yaroslav Tserkovnyak, and Pramey Upadhyaya - \textit{Coherently driving quantum spins via electrically induced non-linear magnetization precessions: a pathway towards high Q-factors} \href{}{\bf (2020)}.
 
\end{etaremune}

\section*{\underline{Invited and Conference Talks}}
\begin{etaremune}
\itemsep-0.1em
\item \textit{Quantum-classical spin hybrids: leveraging spintronic tools for information processing applications}\href{https://doi.org/10.1117/12.2567795}{[Link]}, SPIE Nanoscience + Engineering (2020).

\item \textit{Quantum-Classical Spin Hybrids - a new platform for developing quantum probes and hybrid quantum technologies}, Indiana University-Purdue University Indianapolis, IN (2020).

\item \textit{Coherent Electrical Driving of Quantum Spins via Localized Magnons}, APS March Meeting, Boston, MA (2019).

\item \textit{Room Temperature EHL in monolayer MoS$_2$}, APS March Meeting, Los Angeles, CA (2018).

 \item \textit{Photoemission signature of excitons}, 84$^\mathrm{th}$ annual meeting of SESAPS, Milledgeville, GA (2017).

\item \textit{Non-Equilibrium exciton dynamics in model systems}, APS March Meeting, New Orleans, LA (2017).
 
\item \textit{Coupled Plasmon Phonon Dynamics in GaP: an indirect gap polar semiconductor}, APS March Meeting, Baltimore, MD (2016).

 \item \textit{THz radiation from accelerating photo-excited carriers in graphene}, $5^\mathrm{th}$ International Symposium on Terahertz Nanoscience, Martinique (2015).

\item \textit{Terahertz radiation from accelerating carriers in graphene}, APS March Meeting, Denver, CO (2014).

\item \textit{Noise properties of graphene like systems}, APS March Meeting, Baltimore, MD (2013).
 \end{etaremune}

\vspace{-0.1in}
\section*{\underline{Teaching Experience}}
\begin{itemize}
\item Teaching Assistant, PHY 2053 Lab, Dept. of Physics, Univ. of Florida (2010-2011).
\item Occasional Lectures| ECE 606 at Purdue Univ. (Fall 2018) and PHY 3101 at Univ. of Florida (Fall 2013, Fall 2012).
\item Substitute Teaching Assistant, PHY 2054, Dept. of Physics, Univ. of Florida (Fall 2015).
\end{itemize}

\vspace{-0.1in}
 \section*{\underline{Professional Membership and Services}} 
\textbf{Member}: American Physical Society \\
\textbf{Peer Reviewer}: Nature Communications, Physical Review Letters, Physical Review B, Applied Physics Letters, Optics Communications.
 
% \section*{\underline{References}}
% \begin{enumerate}
 
% \item { \bf Prof. Pramey Upadhyaya} \\
% \email{prameyup@purdue.edu}\\
%  Tel: +1-765-494-5248\\
% School of Electrical and Computer Engineering\\
% Purdue University, West Lafayette, IN\\
 %
% \item { \bf Prof. Alexander F. Kemper} \\
% \email{akemper@ncsu.edu}\\
%  Tel: +1-919-515-7339\\
% Department of Physics\\
% North Carolina State University, Raleigh, NC\\
 %
% \item { \bf Prof. Christopher J. Stanton} \\
% \email{stanton@phys.ufl.edu}\\
% \email{stanton@ufl.edu}\\
%  Tel: +1-352-392-8753\\
% Department of Physics\\
% University of Florida, Gainesville, FL\\
% \end{enumerate}

\end{document}
